A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2019-01-30 paper/submission/ZW10157-response-1.tex
changes on mac that I didn't commit, response and main text

At the same time, AWA is a facility where many different accelerator research experiments take
place. Therefore it is very desirable, and was an important research objective, to develop a precise, e.g. 3D
model embedded into a multi-objective optimization framework. In this way, optimization of various 
machine configurations can be done at different charge levels. While other codes such as GPT  and ELEGANT, also have integrated genetic optimization algorithms; the OPAL framework, differentiates itself by being open source (i.e. free to use), massively parallel, and fully 3D.
}

@@

{\bf Our response:} {\color{blue} The statements on the lines  17-19 of our initial submission are misleading, as
the referee correctly recognized. We replace them with the following:

Despite the success of on-line models in some facilities[ref here], and various model dependent and model independent tuning and optimization techniques, in research facilities like the AWA, empirical tuning by operators is still one of the common methods used. Due to the non-linear problem mainly from space charge, simple and fast models are applicable only in a very restricted manner. When it comes to multi-objective optimization, the situation is even more complicated.
}

@@

@@

@@

{\bf Our response:} {\color{blue} 
	Respectfully, we point out Referee A's comments that 
	"Computing and algorithms" is listed under PRAB's scope. 
	However, for those interested in physics and not implementation details (i.e. user of code not developer), 
	we had hoped the AWA optimization problem would 
	provide a meaningful and tangible use case for the framework.
	It seems we did not present that part of the paper strongly enough. 
	We hope some of the additions above have helped in this area.
	However, would also like to keep some computational details for those who 
	may be interested in code development specifically for accelerator optimization.
}

@@

>>>>

At the same time, AWA is a facility where many different accelerator research experiments take
place. Therefore it is very desirable, and was an important research objective, to develop a precise, e.g. 3D
model embedded into a multi-objective optimization framework. In this way, optimization of various 
machine configurations can be done at different charge levels. While other codes such as GPT [updated ref here] and ELEGANT [updated ref here], also have integrated genetic optimization algorithms; the OPAL framework, differentiates itself by being open source (i.e. free to use), massively parallel, and fully 3D.
}

@@

{\bf Our response:} {\color{blue} The statements on the lines  17-19 of our initial submission are misleading, as
the referee correctly recognized. We replace them with the following on lines 18-27 of the updated manuscript:

Despite the success of on-line models in some facilities [ref here], 
and various model dependent and model independent tuning and optimization techniques, 
empirical tuning by operators a common method used at many research facilities.
When the beam dynamics is non-linear, as is the case with space charge, 
simple and fast models are applicable only in a very restricted manner. 
When it comes to multi-objective optimization, the situation is even more complicated.
In order to be able to reliably identify optimal configurations of
accelerators, we solve large multi-objective design optimization
problems to automate the investigation for an optimal set of tuning parameters.
This approach has been used in the past [updated refs here] with much success.
The difference here being the code implementation and the application to a novel
problem at the Argonne Wakefield Accelerator Facility (AWA).
}

@@

{\bf Our response:} {\color{blue} New references were added, as mentioned above.}

@@

{\bf Our response:} {\color{blue} A line has been here added indicating this capability in GPT too.}

@@

{\bf Our response:} {\color{blue} 
	Respectfully, we point out Referee A's comment that 
	"Computing and algorithms" is listed under PRAB's scope. 
	However, for those interested in physics and not implementation details (i.e. user of code not developer), 
	we had hoped the AWA optimization problem would 
	provide a meaningful and tangible use case for the framework.
	It seems we did not present that part of the paper strongly enough. 
	We hope some of the additions above have helped in this area.
	However, would also like to keep some computational details for those who 
	may be interested in code development specifically for accelerator optimization.
}

@@

{\bf Our Response:} {\color{blue} This optimization was used for design of 
an experiment. There is not an alternative base line, i.e. this is the 
"first such solution for the AWA". They points have not been operationally measured yet.
The following line was added to the abstract:

"Optimized solutions found by the framework were used to finish the design 
of a two beam acceleration experiment."
}

### 2019-01-30 paper/submission/prab.tex
changes on mac that I didn't commit, response and main text

\begin{abstract}
Particle accelerators are invaluable tools for research in the basic and
  applied sciences, in fields such as materials science, chemistry,
  the biosciences, particle physics, nuclear physics and medicine.
The design, commissioning, and operation of accelerator facilities is a
  non-trivial task, due to the large number of control parameters and the
  complex interplay of several conflicting design goals.
We propose to tackle this problem by means of multi-objective optimization
  algorithms which also facilitate a parallel deployment.
In order to compute solutions in a meaningful time frame a fast 
  and scalable software framework is required.
In this paper, we
present a general-purpose framework for simulation-based
multi-objective optimization methods that allows the automatic
investigation of optimal sets of machine parameters.
Using evolutionary algorithms as the optimizer and \textsc{OPAL} as the 
  forward solver, validation experiments and 
 results of multi-objective optimization problems in the domain of beam
  dynamics are presented. The high charge beam line at the Argonne Wakefield 
  Accelerator Facility was used as the beam dynamics model. The 3D beam size, 
  transverse momentum, and energy spread were optimized.

@@

Particle accelerators play a significant role in many aspects of science and
  technology.
Fields, such as material science, chemistry, the biosciences, particle
  physics, nuclear physics and medicine depend on reliable and effective
  particle accelerators, both as research but as well as practical tools.
Achieving the required performance in the design, commissioning, and operation
  of accelerator facilities is a complex and versatile problem.
Today, tuning machine parameters, e.g. bunch charge, emission time and
  various parameters of beamline elements, is most commonly done manually by
  running simulation codes to scan the parameter space.
This approach is tedious, time consuming and can be error prone.
In order to be able to reliably identify optimal configurations of
  accelerators we propose to solve large multi-objective design optimization
  problems to automate the investigation for an optimal set of tuning
  parameters.
Observe that multiple and conflicting optimality criteria call for a
  multi-objective approach.

@@

Optimization problems deal with finding one or more feasible solutions
  corresponding to extreme values of objectives.
If more than one objective is present in the optimization problem this is called
  a multi-objective optimization problem (MOOP).
A MOOP is defined as
%
\begin{align}
  \text{ min} \quad & \quad f_m({\bf x}), ~& m &= 1, \dots, M, \label{eq:moop:obj}\\
  \text{s.t.} \quad & \quad g_j({\bf x}) \geq 0, & j &= 0, \dots, J,
  \label{eq:moop:constr}\\
  \quad & \quad  x_i^L \leq {\bf x}=x_i \leq x_i^U,& i &=0, \dots, n
  \label{eq:moop:dvar} \text{,}
\end{align}
%
where $\bf f$ denotes the objectives (\ref{eq:moop:obj}),
  $\bf g$ the constraints (\ref{eq:moop:constr}),
  and $\bf x$ the design variables (\ref{eq:moop:dvar}).

Often, conflicting objectives are encountered, and this complicates the concept of
  optimality.
This is known as \textit{Pareto optimality}.
The set of Pareto optimal points (blue curve) forms the Pareto front or
  surface.
All points on this surface are considered to be Pareto optimal.

Once the shape of the Pareto front has been determined, the buyer can
  \textit{specify preference}, balancing the features by observing the
  effect on the optimality criteria, converging to the preferred solution.
This is called \textit{a posteriori} preference specification since 
  a solution is selected after all possible trade offs have been presented to us.
An alternative is to specify preference \textit{a priori}, e.g. by
  weighting (specifying preference before solving the problem) and combining
  all objectives into one and applying a single-objective method to solve the
  problem (yielding only one solution).
In many situations preference is not known \textit{a priori} and an
  \textit{a posteriori} preference specification helps conveying a deeper
  understanding of the solution space.
The Pareto front can be explored and the impact of a trade off decision then
  becomes visible.

Sampling Pareto fronts is far from trivial.
A number of approaches have been proposed,
  e.g.\ evolutionary algorithms~\cite{deb:09},
  simulated annealing~\cite{kigv:83},
  swarm methods~\cite{keeb:95},
  and many more~\cite{domc:96,cati:02,kara:05,hoss:09}.
In the next section, we briefly introduce the theory of evolutionary algorithms
  used in the present work.

@@

The task of the variator is to generate offspring and ensure diversity in the
  population.
The variator can start generating offspring once the fitness of every
  individual of the population has been evaluated.
This explicit synchronization point defines an obvious bottleneck for parallel
  implementations of evolutionary algorithms.
In the worst case some MPI processes are taking a long time to compute the
  fitness of the last individual in the pool of individuals to evaluate.
During this time all other resources are idle and wait for the result of
  this one individual in order to continue to generate and evaluate offspring.
To counteract this effect, the selector is already called when two individuals
  have finished evaluating their fitness, lifting the boundaries between
  generations and evaluating the performance of individuals.
New offspring will be generated and MPI processes can immediately return to
  work on the next fitness evaluation.
By calling the selector more frequently (already after two offspring
  individuals have been evaluated) results in better populations since bad
  solutions are rejected earlier.
On the other hand, calling the selector more often is computationally more
  expensive.

Our variator implementation uses the master/slave architecture, presented in
  the next section, to run as many function evaluations as possible in parallel.
Additionally, various crossover and mutation policies are available for tuning
  the algorithm to the optimization problem.

>>>>

\begin{abstract}
Particle accelerators are invaluable tools for research in the basic and
  applied sciences, in fields such as materials science, chemistry,
  the biosciences, particle physics, nuclear physics and medicine.
The design, commissioning, and operation of accelerator facilities is a
  non-trivial task, due to the large number of control parameters and the
  complex interplay of several conflicting design goals.
We propose to tackle this problem by means of multi-objective optimization
  algorithms which also facilitate a parallel deployment.
In order to compute solutions in a meaningful time frame, a fast 
  and scalable software framework is required.
In this paper, we
present a general-purpose framework for simulation-based
multi-objective optimization methods that allows the automatic
investigation of optimal sets of machine parameters.
Using evolutionary algorithms as the optimizer and \textsc{OPAL} as the 
  forward solver, validation experiments and 
 results of multi-objective optimization problems in the domain of beam
  dynamics are presented. The high charge beam line at the Argonne Wakefield 
  Accelerator Facility was used as the beam dynamics model. 
  Optimized solutions found by the framework were used to finish the design 
  of a two beam acceleration experiment.

@@

Particle accelerators play a significant role in many aspects of science and
  technology.
Fields, such as material science, chemistry, the biosciences, particle
  physics, nuclear physics and medicine depend on reliable and effective
  particle accelerators, both as research but as well as practical tools.
Achieving the required performance in the design, commissioning, and operation
  of accelerator facilities is a complex and versatile problem.

Despite the success of on-line models in some facilities [ref here], 
and various model dependent and model independent tuning and optimization techniques, 
empirical tuning by operators a common method used at many research facilities.
When the beam dynamics is non-linear, as is the case with space charge, 
simple and fast models are applicable only in a very restricted manner. 
When it comes to multi-objective optimization, the situation is even more complicated.
In order to be able to reliably identify optimal configurations of
  accelerators, we solve large multi-objective design optimization
  problems to automate the investigation for an optimal set of tuning parameters.
  This approach has been used in the past [] with much success.
The difference here being the implementation and application to a novel
problem at the Argonne Wakefield Accelerator Facility (AWA).

@@

Optimization problems deal with finding one or more feasible solutions
  corresponding to extreme values of objectives.
If more than one objective is present in the optimization problem this is called
  a multi-objective optimization problem (MOOP).
A MOOP is defined as
%
\begin{align}
  \text{ min} \quad & \quad f_m({\bf x}), ~& m &= 1, \dots, M, \label{eq:moop:obj}\\
  \text{s.t.} \quad & \quad g_j({\bf x}) \geq 0, & j &= 0, \dots, J,
  \label{eq:moop:constr}\\
  \quad & \quad  x_i^L \leq {\bf x}=x_i \leq x_i^U,& i &=0, \dots, n
  \label{eq:moop:dvar} \text{,}
\end{align}
%
where $\bf f$ denotes the objectives (\ref{eq:moop:obj}),
  $\bf g$ the constraints (\ref{eq:moop:constr}),
  and $\bf x$ the design variables (\ref{eq:moop:dvar}).
Often, conflicting objectives are encountered, and this complicates the concept of
  optimality. \textit{Pareto optimality} is often used in such situations [].
The set of Pareto optimal points forms the Pareto front or
  surface.
All points on this surface are considered to be Pareto optimal.
Sampling Pareto fronts is far from trivial.
A number of approaches have been proposed,
  e.g.\ evolutionary algorithms~\cite{deb:09},
  simulated annealing~\cite{kigv:83},
  swarm methods~\cite{keeb:95},
  and many more~\cite{domc:96,cati:02,kara:05,hoss:09}.
In the next section, we briefly introduce the theory of evolutionary algorithms
  used in the present work.

@@

The task of the variator is to generate offspring and ensure diversity in the
  population.
The variator can start generating offspring once the fitness of every
  individual of the population has been evaluated.
This explicit synchronization point defines an obvious bottleneck for parallel
  implementations of evolutionary algorithms.
In the worst case, some MPI processes are taking a long time to compute the
  fitness of the last individual in the pool of individuals to evaluate.
During this time all other resources are idle and wait for the result of
  this one individual in order to continue to generate and evaluate offspring.
To counteract this effect, the selector is already called when two individuals
  have finished evaluating their fitness, lifting the boundaries between
  generations and evaluating the performance of individuals.
New offspring will be generated and MPI processes can immediately return to
  work on the next fitness evaluation.
By calling the selector more frequently (already after two offspring
  individuals have been evaluated) results in better populations since bad
  solutions are rejected earlier.
On the other hand, calling the selector more often is computationally more
  expensive. Note this capability is also present in GPT's~\cite{gpt} optimization system.

The variator implementation uses the master/slave architecture, presented in
  the next section, to run as many function evaluations as possible in parallel.
Additionally, various crossover and mutation policies are available for tuning
  the algorithm to the optimization problem.
